**A framework for effective application of machine learning to microbiome-based classification problems**

### 2020-01-28 submission/manuscript.Rmd
begum's edits

We highlighted the importance of model interpretation to gain greater biological insights into microbiota-associated diseases. In this study, we showcased two different interpretation methods: ranking each OTU by (i) their absolute weights in the trained models and (ii) their impact on the predictive performance based on permutation importance. Human-associated microbial communities have complex correlation structures that create collinearity in the datasets. This can hinder our ability to reliably interpret models because the feature weights of correlated OTUs are influenced by one another [@dormann_collinearity:_2013]. To capture all important features, once we identify highly ranked OTUs, we should review their relationships with other OTUs. These relationships will help us generate new hypotheses about the ecology of the disease and test them with follow-up experiments. When we used permutation importance, we partially accounted for collinearity by grouping correlated OTUs to determine their impact as a group. We grouped OTUs that had a perfect correlation with each other; however, we could reduce the correlation threshold to further investigate the relationships among correlated features. By our approach, we identified 432 OTUs out of 6,920 that had perfect correlations with at least one other OTU. The decision to establish correlation thresholds is left to researchers to implement for their own analyses. Regardless of the threshold, undestanding the correlation structures within the data is critical to avoid misinterpreting the models. Such structures are likely to be a particular problem with shotgun metagenomic datasets where collinearity will be more pronounced due to many genes being correlated with one another because they come from the same chromosome. Finally, true causal mechanisms (e.g., role of microbiome in colorectal cancer) cannot be explained solely by the highest performing machine learning model [@li_accurate_2020]. To identify the true underlying microbial factors of a disease, it is crucial to follow up on any correlation analyses with further hypothesis testing and experimentation for biological validation.

>>>>

We highlighted the importance of model interpretation to gain greater biological insights into microbiota-associated diseases. In this study, we showcased two different interpretation methods: ranking each OTU by (i) their absolute weights in the trained models and (ii) their impact on the predictive performance based on permutation importance. Previous studies have emphasized the difficulty of interpreting the feature coefficients in linear models [@breiman_statistical_2001] and the biases introduced by computing feature importance using built-in methods (e.g., gini drop) of tree-based models [@strobl_bias_2007]. Therefore, we encourage our audience to use both interpretation methods highlighted in this study.  Human-associated microbial communities have complex correlation structures that create collinearity in the datasets. This can hinder our ability to reliably interpret models because the feature weights of correlated OTUs are influenced by one another [@dormann_collinearity:_2013]. To capture all important features, once we identify highly ranked OTUs, we should review their relationships with other OTUs. These relationships will help us generate new hypotheses about the ecology of the disease and test them with follow-up experiments. When we used permutation importance, we partially accounted for collinearity by grouping correlated OTUs to determine their impact as a group. We grouped OTUs that had a perfect correlation with each other; however, we could reduce the correlation threshold to further investigate the relationships among correlated features. By our approach, we identified 432 OTUs out of 6,920 that had perfect correlations with at least one other OTU. The decision to establish correlation thresholds is left to researchers to implement for their own analyses. Regardless of the threshold, undestanding the correlation structures within the data is critical to avoid misinterpreting the models. Such structures are likely to be a particular problem with shotgun metagenomic datasets where collinearity will be more pronounced due to many genes being correlated with one another because they come from the same chromosome. Finally, true causal mechanisms (e.g., role of microbiome in colorectal cancer) cannot be explained solely by the highest performing machine learning model [@li_accurate_2020]. To identify the true underlying microbial factors of a disease, it is crucial to follow up on any correlation analyses with further hypothesis testing and experimentation for biological validation.
